---
title: "HW3"
output: html_document
---

```{r}
rm(list = ls())

```

***********************************************************************************
```{r}
library(tidyverse)
library(glmnet)
library(knitr)
library(Lahman)
library(dplyr)

```

```{r}
load('market_level.R') # market-level data is datam
load('market_airline_level.R') # market airline level data
```

Set the seed to 0 and randomly allocate 1,000 rows of the market-level data to a test
set, to be used only in (7). Use the rest to do the following.

```{r}
set.seed(0)
test_rows<- sample(nrow(datam),1000) #randomly choose 1000 rows in mkt-level data
test<- datam[test_rows,] # this test set contains 1000 rows chosen above
df <- datam[-test_rows,] # this df set contains the rest rows

```

```{r}
AA<- #find out which market does American Airline enter
  datama %>%
  filter(ticket_carrier=="AA") %>%
  select(-price:-market_income)

df_AA<-left_join(df,AA, by=c("origin_airport_id"="origin_airport_id",
                             "dest_airport_id"="dest_airport_id"))

df_AA$AA=1*(df_AA$ticket_carrier=="AA") #create an indicator which equals to 1 if AA is in this mkt
df_AA<-select(df_AA,-ticket_carrier,-origin_city:-population_dest)  #drop columns we don't need
df_AA[is.na(df_AA)]<-0 #fill N/A with 0

```


Estimate a linear probability model, predicting whether American Airlines enters a
market as a function of the number of competitors.
```{r}

df_AA<- mutate(df_AA,comp=num_carriers-AA) #create the new variable'numbers of competition'

linear_m<-lm(AA~comp,df_AA)
summary(linear_m)

```

```{r}
#probit model
probit_m<- glm(formula=AA~comp,family = binomial(link='probit'),data = df_AA)
summary(probit_m)

```

```{r}
#logit model
logit_m<- glm(formula=AA~comp,family = binomial(link='logit'),data = df_AA)
summary(logit_m)
```

4) Compute non-parametric estimates of the conditional probabilities of entering. (ie
compute the conditional probability of entering conditional on each number of competitors
directly from the data).

```{r}
num_comp <- sort(unique(df_AA$comp))
non_par<- c()
for (i in num_comp) {
  temp <- filter (df_AA, comp==i)
  prob <- sum(temp$AA==1) /nrow(temp)
  cat('\nThe conditional probability of entntering conditional on ', i,'competitors is :', 
      prob)
  non_par<- rbind(non_par,c(i,prob))
}
 colnames(non_par) <- c('comp', 'prob of AA entering')
 non_par <- as.data.frame(non_par)
 kable(non_par)
```

5)Plot the fitted values of each regression in one graph (i.e. estimated probabilities
on the y-axis and the number of competitors on the x-axis).

```{r}
df_AA<- left_join(df_AA,non_par, by=c('comp'='comp'))
colnames(df_AA)[colnames(df_AA)=="prob of AA entering"] <- "non_par_estimate"
df_AA[is.na(df_AA)]<-0


pred_lm<-predict(linear_m, type='response')
pred_logit<-predict(logit_m, type='response')
pred_probit<-predict(probit_m, type='response')


ggplot(data=df_AA,aes(comp,AA))+
  geom_line(data=df_AA, aes(x=comp, y=non_par_estimate),col = 'green',linetype = 'dashed') +
  #set the non-parametric model to a green dashed line
  geom_line(model=lm, data=df_AA, aes(x=comp, y=pred_lm),col='black')+ 
  # set the linear model to black line
  geom_line(model=glm, data=df_AA, aes(x=comp, y=pred_logit),col='blue')+
  #set the logit model to blue line
  geom_line(model=glm,data=df_AA, aes(x=comp, y=pred_probit),col='red')+
  #set the probit model to red line
  xlab('number of competitors')+ylab('entering')

#a better way to plot: first create a data frame containing predicted outcomes for four models, then plot this data frame, which provides a legend for four models. use gather funtion.
```

In words, explain the coeffi-cients of the first three models. How do the estimated relationships compare? Should we interpret these relationships causally? Are the estimates for the probit and logit similar? Should we have expected this ex ante?
```{r}
coefficients<-c(summary(linear_m)$coefficient[2,1],summary(logit_m)$coefficients[2,1],
                summary(probit_m)$coefficients[2,1])
model<-c('linear','logit','probit')
names(coefficients) <-c(model)
coefficients

```


6  
```{r}
  
mod<-lm(formula=(AA~((comp+ average_distance_m + market_size+ hub_route+ vacation_route+
                      slot_controlled+market_income)^2 + comp^2 + average_distance_m^2+
                       market_size^2 + hub_route^2 + vacation_route^2 + slot_controlled^2-
                       hub_route:slot_controlled)),  x=TRUE ,data = df_AA) 

X<- mod$x[,-1]
y<- mod$model$AA

regular_log<-cv.glmnet(X,y,family='binomial',alpha=1,nfold=10) #create a regularized logistic regression model, using 10 fold

lamda_min<-regular_log$lambda.min #find the optimal lambda
print(lamda_min)
```

7
```{r}
test<-left_join(test,AA, by=c("origin_airport_id"="origin_airport_id",
                             "dest_airport_id"="dest_airport_id")) 
test$AA=1*(test$ticket_carrier=="AA")
test$comp = test$num_carriers-test$AA
test<-select(test,-ticket_carrier,-origin_city:-population_dest)

test<-left_join(test,non_par, by=c('comp'='comp'))
test[is.na(test)]<-0 #fill N/A with 0
```

```{r}
SPE<- function(yhat,y){ #SPE = sum of squared prediction errors
  resd=yhat-y
  SPE=sum(resd^2)/1000
  return(SPE)}

```

```{r}
pre1<-predict(linear_m, newdata = test)
SPE_linear<-SPE(pre1,test$AA)

pre2<-predict(logit_m, newdata = test)
SPE_log<-SPE(pre2,test$AA)

pre3<-predict(probit_m, newdata = test)
SPE_pro<-SPE(pre3,test$AA)

mod2<-lm(formula=(AA~((comp+ average_distance_m + market_size+ hub_route+ vacation_route+
                      slot_controlled+market_income)^2 + comp^2 + average_distance_m^2+
                       market_size^2 + hub_route^2 + vacation_route^2 + slot_controlled^2-
                       hub_route:slot_controlled)),  x=TRUE ,data = test) 


pre4<-predict(regular_log, newx=mod2$x[,-1])
SPE_regular_log <- SPE(pre4,test$AA)

SPE_non_par <- SPE(test$`prob of AA entering`, test$AA)

SPE_table<- matrix(c(SPE_linear,SPE_log,SPE_pro,SPE_regular_log, SPE_non_par),nrow=5,
                   byrow = FALSE)
rownames(SPE_table)<- c('linear model', 'logit model', 'probit model','regularized logit model',
                     'non-parametric model')
colnames(SPE_table) <- c('prediction errors')
kable(SPE_table)

```